---
title: "Identify Index Medications and Export NDC Codes"
author: "Gwen Aubrac"
format: html
---

## Overview

This notebook identifies index medications (GLP-1/GIP agonists and comparators) from the Sentinel drug reference database and exports their NDC codes for use in subsequent programs.

**Index medications:**
- GLP-1/GIP agonists: semaglutide and tirzepatide (liraglutide and dulaglutide are also included in the code list to later exclude prior use of other GLP-1 agonists)
- Comparators: SGLT2 inhibitors, DPP-4 inhibitors, sulfonylureas

**Outputs:**
- Separate CSV files with NDC11 codes for each medication

## Setup

```{r}
#| label: setup
#| message: false

# Load packages
library(tidyverse)
library(here)
library(DBI)
library(odbc)
library(dplyr)
library(dbplyr)
library(stringr)

# Set working directory to wherever this file is for you
project_root <- Sys.getenv("PROJECT_ROOT")
if (nchar(project_root) == 0) stop("PROJECT_ROOT is not set. Please run: source config.sh (bash) or set it in .Renviron (RStudio).")
output_dir  <- file.path(project_root, "output")
results_dir <- file.path(project_root, "novel_res")

# Set up connection to Snowflake
con <- dbConnect(
  odbc::odbc(),
  dsn = Sys.getenv("dsn")
)

# create output directory
if (!dir.exists(output_dir)) {
  dir.create(output_dir)
}

```


## Get NDCs for Index Medications

```{r}
#| label: get-ndcs

DBI::dbExecute(con, "USE DATABASE SENTINEL_COMMON")
DBI::dbExecute(con, "USE SCHEMA SENTINEL_REFERENCE")

drug_reference_tbl <- tbl(
  con,
  "DRUG_REFERENCE"
)

drug_reference_flagged <- drug_reference_tbl %>%
  arrange(PRECISE_GENERIC_NAME, PRODUCT_NAME) %>%
  mutate(

    # Individual GLP-1/GIP medications
    semaglutide = case_when(
      PRECISE_GENERIC_NAME %like% "%SEMAGLUTIDE%" | PRODUCT_NAME %like% "%SEMAGLUTIDE%" ~ 1L,
      PRODUCT_NAME %in% c("SEMAGLUTIDE", "OZEMPIC", "RYBELSUS", "WEGOVY") ~ 1L,
      TRUE ~ 0L
    ),
    tirzepatide = case_when(
      PRECISE_GENERIC_NAME %like% "%TIRZEPATIDE%" | PRODUCT_NAME %like% "%TIRZEPATIDE%" ~ 1L,
      PRODUCT_NAME %in% c("TIRZEPATIDE", "MOUNJARO", "ZEPBOUND") ~ 1L,
      TRUE ~ 0L
    ),
    liraglutide = case_when(
      PRECISE_GENERIC_NAME %like% "%LIRAGLUTIDE%" | PRODUCT_NAME %like% "%LIRAGLUTIDE%" ~ 1L,
      PRODUCT_NAME %in% c("LIRAGLUTIDE", "VICTOZA", "SAXENDA") ~ 1L,
      TRUE ~ 0L
    ),
    dulaglutide = case_when(
      PRECISE_GENERIC_NAME %like% "%DULAGLUTIDE%" | PRODUCT_NAME %like% "%DULAGLUTIDE%" ~ 1L,
      PRODUCT_NAME %in% c("DULAGLUTIDE", "TRULICITY") ~ 1L,
      TRUE ~ 0L
    ),

    # GLP-1/GIP agonists as a class (combine individual medications)
    glp1 = case_when(
      semaglutide == 1L | tirzepatide == 1L | liraglutide == 1L | dulaglutide == 1L ~ 1L,
      TRUE ~ 0L
    ),

    # SGLT2 inhibitors (comparator)
    sglt2 = case_when(
      PRECISE_GENERIC_NAME %like% "%BEXAGLIFLOZIN%" | PRODUCT_NAME %like% "%BEXAGLIFLOZIN%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%CANAGLIFLOZIN%" | PRODUCT_NAME %like% "%CANAGLIFLOZIN%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%DAPAGLIFLOZIN%" | PRODUCT_NAME %like% "%DAPAGLIFLOZIN%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%EMPAGLIFLOZIN%" | PRODUCT_NAME %like% "%EMPAGLIFLOZIN%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%ERTUGLIFLOZIN%" | PRODUCT_NAME %like% "%ERTUGLIFLOZIN%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%SOTAGLIFLOZIN%" | PRODUCT_NAME %like% "%SOTAGLIFLOZIN%" ~ 1L,
      PRODUCT_NAME %in% c("BEXAGLIFLOZIN", "CANAGLIFLOZIN", "DAPAGLIFLOZIN", "EMPAGLIFLOZIN", 
                          "ERTUGLIFLOZIN", "SOTAGLIFLOZIN",
                          "BRENZAVVY", "INVOKANA", "FARXIGA", "JARDIANCE", "STEGLATRO", "INPEFA") ~ 1L,
      TRUE ~ 0L
    ),

    # DPP-4 inhibitors (comparator)
    dpp4 = case_when(
      PRECISE_GENERIC_NAME %like% "%SITAGLIPTIN%" | PRODUCT_NAME %like% "%SITAGLIPTIN%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%SAXAGLIPTIN%" | PRODUCT_NAME %like% "%SAXAGLIPTIN%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%LINAGLIPTIN%" | PRODUCT_NAME %like% "%LINAGLIPTIN%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%ALOGLIPTIN%" | PRODUCT_NAME %like% "%ALOGLIPTIN%" ~ 1L,
      PRODUCT_NAME %in% c("SITAGLIPTIN", "SAXAGLIPTIN", "LINAGLIPTIN", "ALOGLIPTIN",
                          "JANUVIA", "ONGLYZA", "TRADJENTA", "NESINA") ~ 1L,
      TRUE ~ 0L
    ),

    # Sulfonylureas (comparator)
    sulfonylurea = case_when(
      PRECISE_GENERIC_NAME %like% "%GLIPIZIDE%" | PRODUCT_NAME %like% "%GLIPIZIDE%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%GLYBURIDE%" | PRODUCT_NAME %like% "%GLYBURIDE%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%GLIMEPIRIDE%" | PRODUCT_NAME %like% "%GLIMEPIRIDE%" ~ 1L,
      PRECISE_GENERIC_NAME %like% "%GLICLAZIDE%" | PRODUCT_NAME %like% "%GLICLAZIDE%" ~ 1L,
      PRODUCT_NAME %in% c("GLIPIZIDE", "GLYBURIDE", "GLIMEPIRIDE", "GLICLAZIDE",
                          "GLUCOTROL", "DIABETA", "MICRONASE", "AMARYL", "DIAMICRON") ~ 1L,
      TRUE ~ 0L
    )
  )

# Collect the drug reference data
drug_reference <- drug_reference_flagged %>%
  select(CODE, PRECISE_GENERIC_NAME, PRODUCT_NAME, KH_THERAPEUTIC_CLASS_ARRAY, 
         glp1, semaglutide, tirzepatide, liraglutide, dulaglutide,
         sglt2, dpp4, sulfonylurea) %>%
  filter(glp1 == 1 | sglt2 == 1 | dpp4 == 1 | sulfonylurea == 1) %>%
  collect()

# Check counts for each medication class
drug_reference %>%
  summarise(
    glp1_count = sum(glp1),
    semaglutide_count = sum(semaglutide),
    tirzepatide_count = sum(tirzepatide),
    liraglutide_count = sum(liraglutide),
    dulaglutide_count = sum(dulaglutide),
    sglt2_count = sum(sglt2),
    dpp4_count = sum(dpp4),
    sulfonylurea_count = sum(sulfonylurea)
  )

# Preview medications by class
drug_reference %>%
  filter(glp1 == 1) %>%
  select(CODE, PRECISE_GENERIC_NAME, PRODUCT_NAME, KH_THERAPEUTIC_CLASS_ARRAY) %>%
  arrange(PRECISE_GENERIC_NAME)

```

## Manually Remove NDC Codes that Correspond to Multiple Classes

This step removes NDC codes that correspond to multiple drug classes (more details on this later). You can go back to this step to add more medications as needed. 

```{r}
#| label: remove-multi-ndcs

drug_reference <- drug_reference %>%
  filter(PRODUCT_NAME != "GLYXAMBI" & PRODUCT_NAME != "QTERN" & PRODUCT_NAME != "STEGLUJAN" & PRODUCT_NAME != "TRIJARDY XR")
```

## Export NDC Codes

```{r}
#| label: export-ndcs

# Export NDC codes separately for each medication class and individual GLP-1 medications

# Individual GLP-1 medications
drug_reference %>%
  filter(semaglutide == 1) %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  select(
    NDC11,
    GENERIC_NAME = PRECISE_GENERIC_NAME,
    PRODUCT_NAME,
    THERAPEUTIC_CLASS = KH_THERAPEUTIC_CLASS_ARRAY
  ) %>%
  write_csv(file.path(output_dir, "semaglutide_ndcs.csv"))

drug_reference %>%
  filter(tirzepatide == 1) %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  select(
    NDC11,
    GENERIC_NAME = PRECISE_GENERIC_NAME,
    PRODUCT_NAME,
    THERAPEUTIC_CLASS = KH_THERAPEUTIC_CLASS_ARRAY
  ) %>%
  write_csv(file.path(output_dir, "tirzepatide_ndcs.csv"))

drug_reference %>%
  filter(liraglutide == 1) %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  select(
    NDC11,
    GENERIC_NAME = PRECISE_GENERIC_NAME,
    PRODUCT_NAME,
    THERAPEUTIC_CLASS = KH_THERAPEUTIC_CLASS_ARRAY
  ) %>%
  write_csv(file.path(output_dir, "liraglutide_ndcs.csv"))

drug_reference %>%
  filter(dulaglutide == 1) %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  select(
    NDC11,
    GENERIC_NAME = PRECISE_GENERIC_NAME,
    PRODUCT_NAME,
    THERAPEUTIC_CLASS = KH_THERAPEUTIC_CLASS_ARRAY
  ) %>%
  write_csv(file.path(output_dir, "dulaglutide_ndcs.csv"))

# All GLP-1 medications combined
drug_reference %>%
  filter(glp1 == 1) %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  select(
    NDC11,
    GENERIC_NAME = PRECISE_GENERIC_NAME,
    PRODUCT_NAME,
    THERAPEUTIC_CLASS = KH_THERAPEUTIC_CLASS_ARRAY
  ) %>%
  write_csv(file.path(output_dir, "glp1_ndcs.csv"))

# Comparator classes
drug_reference %>%
  filter(sglt2 == 1) %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  select(
    NDC11,
    GENERIC_NAME = PRECISE_GENERIC_NAME,
    PRODUCT_NAME,
    THERAPEUTIC_CLASS = KH_THERAPEUTIC_CLASS_ARRAY
  ) %>%
  write_csv(file.path(output_dir, "sglt2_ndcs.csv"))

drug_reference %>%
  filter(dpp4 == 1) %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  select(
    NDC11,
    GENERIC_NAME = PRECISE_GENERIC_NAME,
    PRODUCT_NAME,
    THERAPEUTIC_CLASS = KH_THERAPEUTIC_CLASS_ARRAY
  ) %>%
  write_csv(file.path(output_dir, "dpp4_ndcs.csv"))

drug_reference %>%
  filter(sulfonylurea == 1) %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  select(
    NDC11,
    GENERIC_NAME = PRECISE_GENERIC_NAME,
    PRODUCT_NAME,
    THERAPEUTIC_CLASS = KH_THERAPEUTIC_CLASS_ARRAY
  ) %>%
  write_csv(file.path(output_dir, "sulfonylurea_ndcs.csv"))
```

Here we check whether there are NDC codes that correspond to multiple drug classes. If any are found, there will be a warning and you can check which NDC codes, and product names were flagged in the "ndc_overlap_summary.txt". Please check the log to see if there are any NDCs that correspond to multiple medications. If so, please exclude them from the NDC lists, which will later be used to identify patients for inclusion in the cohort. 

```{r}
# Check for duplicate NDC11 codes across medication classes
ndc_overlap <- drug_reference %>%
  mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
  group_by(NDC11) %>%
  summarise(
    n_classes = sum(glp1) + sum(sglt2) + sum(dpp4) + sum(sulfonylurea),
    in_glp1 = max(glp1),
    in_sglt2 = max(sglt2),
    in_dpp4 = max(dpp4),
    in_sulfonylurea = max(sulfonylurea),
    .groups = "drop"
  ) %>%
  filter(n_classes > 1)

if (nrow(ndc_overlap) > 0) {
  warning("Found ", nrow(ndc_overlap), " NDC codes assigned to multiple medication classes")
  
  # Get detailed information about overlapping NDCs
  overlap_details <- drug_reference %>%
    mutate(NDC11 = str_pad(as.character(CODE), width = 11, pad = "0")) %>%
    filter(NDC11 %in% ndc_overlap$NDC11) %>%
    select(NDC11, PRODUCT_NAME, PRECISE_GENERIC_NAME,
           semaglutide, tirzepatide, liraglutide, dulaglutide,
           sglt2, dpp4, sulfonylurea) %>%
    pivot_longer(
      cols = c(semaglutide, tirzepatide, liraglutide, dulaglutide, sglt2, dpp4, sulfonylurea),
      names_to = "medication_class",
      values_to = "flag"
    ) %>%
    filter(flag == 1) %>%
    group_by(NDC11, PRODUCT_NAME, PRECISE_GENERIC_NAME) %>%
    summarise(
      classes = paste(medication_class, collapse = ", "),
      n_classes = n(),
      .groups = "drop"
    ) %>%
    arrange(desc(n_classes), PRODUCT_NAME)
  
  # Save detailed overlap information
  write_csv(overlap_details, file.path(output_dir, "ndc_overlap_details.csv"))
  
  summary_text <- c(
    "=======================================================",
    "NDC Overlap:",
    "=======================================================",
    paste("\nDate:", Sys.Date()),
    paste("Total NDC codes that belong to multiple classes:", nrow(overlap_details))
  )
  
  for (i in 1:nrow(overlap_details)) {
    product_line <- paste0(
      "- NDC: ", overlap_details$NDC11[i], " | ",
      overlap_details$PRODUCT_NAME[i],
      " (", overlap_details$classes[i], ")"
    )
    summary_text <- c(summary_text, product_line)
  }
  
  # Write summary to text file
  writeLines(summary_text, file.path(output_dir, "ndc_overlap_summary.txt"))
  
  # Display summary in notebook
  cat(paste(summary_text, collapse = "\n"))
  
} else {
  message("No NDC codes overlap across medication classes")
}

```