---
title: "Extract Index Medication Dispensings"
author: "Gwen Aubrac"
format: html
---

## Overview

This notebook extracts pharmacy claims for index medications from Snowflake, creating chronological dispensing histories for each medication. We are getting the dispensings for one medication of interest at a time. We end up with all of the pharmacy dispensings for the drugs of interest available in the data, which will later be used to identify date of first use (new users) during the study periods of interest and also treatment adherence over time. 

## Setup

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(here)
library(DBI)
library(odbc)
library(dplyr)
library(dbplyr)
library(arrow)

project_root <- Sys.getenv("PROJECT_ROOT")
if (nchar(project_root) == 0) stop("PROJECT_ROOT is not set. Please run: source config.sh (bash) or set it in .Renviron (RStudio).")
output_dir  <- file.path(project_root, "output")
results_dir <- Sys.getenv("RESULTS_DIR", unset = file.path(project_root, "results"))

# Set up connection to Snowflake
con <- dbConnect(
  odbc::odbc(),
  dsn = Sys.getenv("dsn")
)

# Initialize log file for extraction summary
log_lines <- c(
  "=======================================================",
  "Pharmacy Dispensing Extraction Summary",
  "=======================================================",
  paste("Date:", Sys.Date()),
  paste("Time:", format(Sys.time(), "%H:%M:%S")),
  ""
)

```

## Check NDC Format and Set Up Filters

```{r}
#| label: verify-ndc-format
#| message: false

# Set Snowflake context to pharmacy data
DBI::dbExecute(con, "USE DATABASE DSVC_RWJF_BU_AA_RE_ENCOUNTERS_PROD")
DBI::dbExecute(con, "USE SCHEMA COHORT_1302462")

# Load pharmacy claims table (read-only)
pharmacy_tbl <- tbl(con, "PHARMACY_LATEST")

# Check NDC11 format in PHARMACY_LATEST
# to make sure the format is the same in PHARMACY_LATEST and our NDC code files
pharmacy_sample <- pharmacy_tbl %>%
  select(NDC11) %>%
  head(10) %>%
  collect()

pharmacy_sample

# Check NDC11 format in  exported files
ndc_sample <- read_csv(file.path(output_dir, "semaglutide_ndcs.csv"), show_col_types = FALSE) %>%
  head(5) %>%
  pull(NDC11)

ndc_sample

# Select just the columns we are interested in
pharmacy_filtered <- pharmacy_tbl %>%
  select(
    PATIENT_ID,
    DATE_OF_SERVICE,
    DAYS_SUPPLY,
    PRESCRIBER_NPI,
    NDC11,
    TRANSACTION_TYPE
  )

```

## Extract and Save Dispensings

To do this, we are writing a function that:
1. Pulls the pharmacy data from Komodo and filters for NDC codes in the list we provide
2. Create a "med_class" variable to keep track of which medication class we're pulling the fills for (for GLP1s we will also have glp1_med to keep track of which specific GLP1 medication it is))
3. Keep track of number of fills found, number of unique patients, and average dispensings per patient
4. Save as a compresset parquet file and return the information (to the log and the console)

Some filters we are applying to the dispensings:
- Only PAID transactions
- Only valid days fill (≠0)

```{r}
#| label: extract-dispensings
#| message: false

# Function to extract, clean, and save dispensings for one medication at a time
# This minimizes memory usage by not holding multiple large datasets in memory
extract_and_save_dispensings <- function(ndc_file, medication_name, med_class, 
                                         output_file, glp1_med = NA) {
  
  message("\n========================================")
  message("Processing: ", medication_name)
  message("========================================")
  
  # Load NDC list
  ndc_list <- read_csv(ndc_file, show_col_types = FALSE) %>%
    pull(NDC11)
  
  message("  Found ", length(ndc_list), " unique NDCs")
  
  # Join with pharmacy claims, collect, clean, and save in one pipeline
  # and filter pharmacy claims for paid dispensings with valid days supply
  dispensings_temp <- pharmacy_filtered %>%
    filter(NDC11 %in% ndc_list) %>%
    filter(TRANSACTION_TYPE == "PAID") %>%
    filter(DAYS_SUPPLY != 0) %>%
    collect() %>%
    mutate(
      DATE_OF_SERVICE = as.Date(DATE_OF_SERVICE),
      med_class = med_class,
      glp1_med = glp1_med
    ) %>%
    arrange(PATIENT_ID, DATE_OF_SERVICE) %>%
    select(-NDC11)  # Remove NDC11 to reduce table size
  
  # Get statistics before saving
  n_records <- nrow(dispensings_temp)
  n_patients <- n_distinct(dispensings_temp$PATIENT_ID)
  dispensings_per_patient <- round(n_records / n_patients, 1)
  
  message("  Retrieved ", format(n_records, big.mark = ","), " dispensing records")
  message("  Unique patients: ", format(n_patients, big.mark = ","))
  message("  Dispensings per patient: ", dispensings_per_patient)
  
  # Save as compressed Parquet file
  write_parquet(dispensings_temp, output_file)
  message("  Saved to: ", output_file)
  
  # Add to log
  log_lines <<- c(
    log_lines,
    paste0("Medication: ", medication_name),
    paste0("  NDC codes: ", length(ndc_list)),
    paste0("  Dispensing records: ", format(n_records, big.mark = ",")),
    paste0("  Unique patients: ", format(n_patients, big.mark = ",")),
    paste0("  Dispensings per patient: ", dispensings_per_patient),
    paste0("  Output file: ", output_file),
    ""
  )
  
  # Return statistics for summary table
  stats <- list(
    medication = medication_name,
    n_dispensings = n_records,
    n_patients = n_patients,
    dispensings_per_patient = dispensings_per_patient
  )
  
  # Clean up temp object and force garbage collection
  rm(dispensings_temp)
  gc()
  
  return(stats)
}

# Process each medication one at a time to minimize memory usage
stats_semaglutide <- extract_and_save_dispensings(
  file.path(output_dir, "semaglutide_ndcs.csv"), # NDC code file
  "Semaglutide", # medication name (medication_name)
  "GLP1", # medication group name (med_class)
  file.path(output_dir, "dispensings_semaglutide.parquet"), # where to save the file
  "semaglutide" # GLP1 medication name if applicable (glp1_med)
)

stats_tirzepatide <- extract_and_save_dispensings(
  file.path(output_dir, "tirzepatide_ndcs.csv"),
  "Tirzepatide",
  "GLP1",
  file.path(output_dir, "dispensings_tirzepatide.parquet"),
  "tirzepatide"
)

stats_liraglutide <- extract_and_save_dispensings(
  file.path(output_dir, "liraglutide_ndcs.csv"),
  "Liraglutide",
  "GLP1",
  file.path(output_dir, "dispensings_liraglutide.parquet"),
  "liraglutide"
)

stats_dulaglutide <- extract_and_save_dispensings(
  file.path(output_dir, "dulaglutide_ndcs.csv"),
  "Dulaglutide",
  "GLP1",
  file.path(output_dir, "dispensings_dulaglutide.parquet"),
  "dulaglutide"
)

stats_sglt2 <- extract_and_save_dispensings(
  file.path(output_dir, "sglt2_ndcs.csv"),
  "SGLT2i",
  "SGLT2",
  file.path(output_dir, "dispensings_sglt2.parquet")
)

stats_dpp4 <- extract_and_save_dispensings(
  file.path(output_dir, "dpp4_ndcs.csv"),
  "DPP-4i",
  "DPP4",
  file.path(output_dir, "dispensings_dpp4.parquet")
)

stats_sulfo <- extract_and_save_dispensings(
  file.path(output_dir, "sulfonylurea_ndcs.csv"),
  "Sulfonylureas",
  "SULFO",
  file.path(output_dir, "dispensings_sulfo.parquet")
)

```

## Create Combined GLP-1 Dataset

Combine all the fills for the GLP1 medications (arranged by ID and date of service) and similarly compute summary statistics. 

```{r}
#| label: create-glp1-combined

# Load individual GLP-1 files and combine (more memory efficient than keeping in memory)
dispensings_glp1 <- bind_rows(
  read_parquet(file.path(output_dir, "dispensings_semaglutide.parquet")),
  read_parquet(file.path(output_dir, "dispensings_tirzepatide.parquet")),
  read_parquet(file.path(output_dir, "dispensings_liraglutide.parquet")),
  read_parquet(file.path(output_dir, "dispensings_dulaglutide.parquet"))
) %>%
  arrange(PATIENT_ID, DATE_OF_SERVICE)

message("  Combined GLP-1 dispensings: ", format(nrow(dispensings_glp1), big.mark = ","), " records")
message("  Unique GLP-1 patients: ", format(n_distinct(dispensings_glp1$PATIENT_ID), big.mark = ","))

# Save combined dataset
write_parquet(dispensings_glp1, file.path(output_dir, "dispensings_glp1.parquet"))
message("  Saved to: ../output/dispensings_glp1.parquet")

# Calculate stats for log
stats_glp1_all <- list(
  medication = "All GLP-1",
  n_dispensings = nrow(dispensings_glp1),
  n_patients = n_distinct(dispensings_glp1$PATIENT_ID),
  dispensings_per_patient = round(nrow(dispensings_glp1) / n_distinct(dispensings_glp1$PATIENT_ID), 1)
)

log_lines <- c(
  log_lines,
  "Combined GLP-1 Dataset:",
  paste0("  Total dispensing records: ", format(stats_glp1_all$n_dispensings, big.mark = ",")),
  paste0("  Unique patients: ", format(stats_glp1_all$n_patients, big.mark = ",")),
  paste0("  Dispensings per patient: ", stats_glp1_all$dispensings_per_patient),
  paste0("  Output file: ../output/dispensings_glp1.parquet"),
  ""
)

# Clean up
gc()

```

## Summary Statistics

```{r}
#| label: summary-stats

# Create summary table from collected statistics
dispensing_summary <- bind_rows(
  stats_semaglutide,
  stats_tirzepatide,
  stats_liraglutide,
  stats_dulaglutide,
  stats_glp1_all,
  stats_sglt2,
  stats_dpp4,
  stats_sulfo
)

print(dispensing_summary)

# Add summary table to log
log_lines <- c(
  log_lines,
  "Summary Table:",
  ""
)

for (i in 1:nrow(dispensing_summary)) {
  log_lines <- c(
    log_lines,
    paste0(dispensing_summary$medication[i], ":"),
    paste0("  Dispensings: ", format(dispensing_summary$n_dispensings[i], big.mark = ",")),
    paste0("  Patients: ", format(dispensing_summary$n_patients[i], big.mark = ",")),
    paste0("  Dispensings/patient: ", dispensing_summary$dispensings_per_patient[i]),
    ""
  )
}

# Write log file

writeLines(log_lines, file.path(output_dir, "dispensing_extraction_log.txt"))

# Disconnect from Snowflake
dbDisconnect(con)

```


# Save Dispensing Files to Stata

I originally tried to get the discontinuation dates using R, but it was just not possible due to the file size. Could not figure out how to optimize in R, but Stata on the other hand was very efficient. So I converted the files to Stata format. For space, we will delete them after getting the censoring dates later. 

```{r}
#|label: convert-stata
library(haven)

dispensings_glp1 <- read_parquet(file.path(output_dir, "dispensings_glp1.parquet"))
write_dta(dispensings_glp1, file.path(output_dir, "dispensings_glp1.dta"))
rm(dispensings_glp1)

dispensings_sglt2 <- read_parquet(file.path(output_dir, "dispensings_sglt2.parquet"))
write_dta(dispensings_sglt2, file.path(output_dir, "dispensings_sglt2.dta"))
rm(dispensings_sglt2)

dispensings_dpp4 <- read_parquet(file.path(output_dir, "dispensings_dpp4.parquet"))
write_dta(dispensings_dpp4, file.path(output_dir, "dispensings_dpp4.dta"))
rm(dispensings_dpp4)

dispensings_sulfo <- read_parquet(file.path(output_dir, "dispensings_sulfo.parquet"))
write_dta(dispensings_sulfo, file.path(output_dir, "dispensings_sulfo.dta"))
rm(dispensings_sulfo)

dispensings_semaglutide <- read_parquet(file.path(output_dir, "dispensings_semaglutide.parquet"))
write_dta(dispensings_semaglutide, file.path(output_dir, "dispensings_semaglutide.dta"))
rm(dispensings_semaglutide)

dispensings_tirzepatide <- read_parquet(file.path(output_dir, "dispensings_tirzepatide.parquet"))
write_dta(dispensings_tirzepatide, file.path(output_dir, "dispensings_tirzepatide.dta"))
rm(dispensings_tirzepatide)

dispensings_liraglutide <- read_parquet(file.path(output_dir, "dispensings_liraglutide.parquet"))
write_dta(dispensings_liraglutide, file.path(output_dir, "dispensings_liraglutide.dta"))
rm(dispensings_liraglutide)

dispensings_dulaglutide <- read_parquet(file.path(output_dir, "dispensings_dulaglutide.parquet"))
write_dta(dispensings_dulaglutide, file.path(output_dir, "dispensings_dulaglutide.dta"))
rm(dispensings_dulaglutide)
```