---
title: "IPTW Model - Novel GLP1s vs other classes"
author: "Gwen Aubrac"
format: html
embed-resources: true
---

## Overview

The emulated trials are as follows:
- "data1_novel": novel GLP1s vs SGLT2
- "data2_novel": novel GLP1s vs DPP4
- "data3_novel": novel GLP1s vs sulfonylureas

Save results in /../results

Weights are truncated at the 2.5th and 97.5th percentiles. Entropy balancing weights are calculated for a sensitivity analysis at the end and saved as "ebal".

**Sensitivity-controlled options** (set by `SCENARIO` — see `sensitivity.R`):
- `weight_trimming` + `trimming_pct`: when TRUE, drop patients with propensity scores outside the asymmetric quantile bounds (comparator upper `1 - trimming_pct`, treated lower `trimming_pct`) before computing IPTW.

Entropy balancing weights are computed at the end and saved as `ebal` (the `weight_method = "ebal"` scenario uses them in the survival analyses).

Note: trimming before or after computing the weights does not matter.

```{r}
#| label: setup
#### IPTW USING BASELINE COVARIATES ####

library(dplyr)
library(magrittr)
library(haven)
library(tidyr)

project_root <- Sys.getenv("PROJECT_ROOT")
if (nchar(project_root) == 0) stop("PROJECT_ROOT is not set. Launch the pipeline via run_all.R, source config.R in your R session, or set it in .Renviron.")
output_dir  <- file.path(project_root, "output")
results_dir <- Sys.getenv("RESULTS_DIR", unset = file.path(project_root, "results"))

source(file.path(project_root, "sensitivity.R"))
SCENARIO <- get_scenario()

data <- readRDS(file.path(output_dir, "ipcw_novel.rds"))
cov_list <- readRDS(file.path(output_dir, "cov_list_novel.rds"))

data %<>%
    mutate(
        novel_flag = if_else(index_glp1 == "semaglutide" | index_glp1 == "tirzepatide", 1, 0)
    )
```

```{r}
#| label: iptw-by-cohort
# data1: sglt2 (ref) vs novel GLP1s (comp)
data1 <- data %>%
  filter(novel_flag == 1 | index_class == "SGLT2") %>%
  mutate(trt = as.factor(if_else(novel_flag == 1, 1, 0))) # those with novel GLP1s are "treated"

summary(data1)

ps_mod_crude1 <- glm(trt ~ 1, family = binomial(), data = data1)

ps_formula <- as.formula(paste("trt ~", paste(cov_list, collapse = " + ")))
ps_formula

ps_mod_adj1 <- glm(
  ps_formula,
  data = data1, 
  family = binomial()
)

data1$ps <- predict(ps_mod_adj1, type = "response", newdata = data1)
data1$ps_marg <- predict(ps_mod_crude1, type = "response")

# Sensitivity: asymmetric PS trimming at trimming_pct / (1 - trimming_pct).
if (isTRUE(SCENARIO$weight_trimming)) {
  .upper_q <- 1 - SCENARIO$trimming_pct
  .lower_q <- SCENARIO$trimming_pct
  perc_upper <- data1 %>% filter(trt == 0) %>%
    summarise(perc_upper = quantile(ps, .upper_q, na.rm = TRUE))
  perc_lower <- data1 %>% filter(trt == 1) %>%
    summarise(perc_lower = quantile(ps, .lower_q, na.rm = TRUE))
  data1 %<>%
    mutate(
      perc_upper = round(perc_upper$perc_upper, 4),
      perc_lower = round(perc_lower$perc_lower, 4)
    )
  data1 <- data1 %>% filter(ps < perc_upper & ps > perc_lower)
  rm(perc_upper, perc_lower)
}

data1$iptw <- if_else(data1$trt ==1, 1/data1$ps, 1/(1-data1$ps))
data1$siptw <- data1$iptw * data1$ps_marg    

# check which patients have extreme weights
summary(data1$siptw)
test <- data1 %>%
  filter(siptw > 30)

summary(test)

# truncate at 2.5 and 97.5 percentiles
data1$iptw_notrunc <- data1$iptw
data1$siptw_notrunc <- data1$siptw

lower_bound <- quantile(data1$siptw, 0.025)
upper_bound <- quantile(data1$siptw, 0.975)
data1$siptw <- pmin(pmax(data1$siptw, lower_bound), upper_bound)
rm(lower_bound, upper_bound)

lower_bound <- quantile(data1$iptw, 0.025)
upper_bound <- quantile(data1$iptw, 0.975)
data1$iptw <- pmin(pmax(data1$iptw, lower_bound), upper_bound)

summary(data1$iptw)
summary(data1$siptw)

# check balance and update ps model accordingly
library(cobalt)
weights_out1 <- WeightIt::weightit(
  formula = ps_formula,
  data = data1,
  method = "glm"
)

# truncate at 2.5 and 97.5 percentiles
lower <- quantile(weights_out1$weights, 0.025)
upper <- quantile(weights_out1$weights, 0.975)
weights_out1$weights <- pmin(pmax(weights_out1$weights, lower), upper)

summary(weights_out1$weights)


weights_out1$treat <- if_else(weights_out1$treat == 1, "Novel GLP1", "SGLT2")

library(ggplot2)

p <- bal.plot(
  weights_out1,
  var.name = "prop.score",
  which = "both",
  type = "density",
  mirror = FALSE,
  grid = FALSE,
  legend.labels = c("Novel GLP1", "SGLT2"),
  position = "bottom"
)

p <- p + 
  ggtitle("Propensity Score Distributional Balance") +
  xlab("Propensity score") +
  ylab("Density")

p

ggsave(file.path(results_dir, "ps_bal1.png"), plot = p, width = 6, height = 4, units = "in", bg = "white")
dev.off()
rm(p)

bal_tab1 <- bal.tab(ps_formula,
                    data = data1,
                    weights = 'iptw',
                    binary = 'std',
                    continuous = 'std',
                    stats = c('mean.diffs'),
                    s.d.denom = 'pooled',
                    un = TRUE,
                    thresholds = c(m=0.1)
                    )

bal1 <- bal_tab1$Balance %>%
  dplyr::rename('SMD (Unadjusted)' = Diff.Un,
                'SMD (IPTW)' = Diff.Adj,
                'Balance' = M.Threshold)

bal1$Variable <- row.names(bal1)
write.table(bal1, file.path(results_dir, "bal1.csv"), col.names = T, row.names = T, append = F, sep = ',')

p <- love.plot(
  bal_tab1, 
  binary = "std",
  thresholds = c(m = .1),
  title = "Covariate Balance with IPTW",
  position = "bottom",
  var.order = "unadjusted",
  size = 2
)

p <- p + theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color = "black"),
  
  axis.text.x = element_text(size = 8),
  axis.text.y = element_text(size = 8),
  axis.title.x = element_text(size = 8, face = 'bold'),
  plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
  
  legend.position = "bottom", 
  legend.title = element_text(size = 8, face = 'bold'),
  legend.text = element_text(size = 8),
  legend.box.margin = margin(0, 0, 0, 0),
  legend.margin = margin(0, 0, 0, 0),
  legend.key.size = unit(0.5, "lines")
) 

p

ggsave(file.path(results_dir, "loveplot1.png"), plot = p, width = 6, height = 12, units = "in", bg = 'white')
dev.off()
rm(p)

# repeat for other comparisons:
# data2: dpp4 (ref) vs novel GLP1s (comp)
data2 <- data %>%
  filter(novel_flag == 1 | index_class == "DPP4") %>%
  mutate(trt = as.factor(if_else(novel_flag == 1, 1, 0)))

ps_mod_crude2 <- glm(trt ~ 1, data = data2, family = binomial)

ps_mod_adj2 <- glm(
  ps_formula,
  family = binomial,
  data = data2
)

data2$ps <- predict(ps_mod_adj2, type = "response")
data2$ps_marg <- predict(ps_mod_crude2, type = "response")

# Sensitivity: asymmetric PS trimming at trimming_pct / (1 - trimming_pct).
if (isTRUE(SCENARIO$weight_trimming)) {
  .upper_q <- 1 - SCENARIO$trimming_pct
  .lower_q <- SCENARIO$trimming_pct
  perc_upper <- data2 %>% filter(trt == 0) %>%
    summarise(perc_upper = quantile(ps, .upper_q, na.rm = TRUE))
  perc_lower <- data2 %>% filter(trt == 1) %>%
    summarise(perc_lower = quantile(ps, .lower_q, na.rm = TRUE))
  data2 %<>%
    mutate(
      perc_upper = round(perc_upper$perc_upper, 4),
      perc_lower = round(perc_lower$perc_lower, 4)
    )
  data2 <- data2 %>% filter(ps < perc_upper & ps > perc_lower)
  rm(perc_upper, perc_lower)
}

data2$iptw <- if_else(data2$trt ==1, 1/data2$ps, 1/(1-data2$ps))
data2$siptw <- data2$iptw * data2$ps_marg    

summary(data1$siptw)
test <- data1 %>%
  filter(siptw > 30)

summary(test)

# truncate at 2.5 and 97.5 percentiles
data2$iptw_notrunc <- data2$iptw
data2$siptw_notrunc <- data2$siptw

lower_bound <- quantile(data2$siptw, 0.025)
upper_bound <- quantile(data2$siptw, 0.975)
data2$siptw <- pmin(pmax(data2$siptw, lower_bound), upper_bound)
rm(lower_bound, upper_bound)

lower_bound <- quantile(data2$iptw, 0.025)
upper_bound <- quantile(data2$iptw, 0.975)
data2$iptw <- pmin(pmax(data2$iptw, lower_bound), upper_bound)

weights_out2 <- WeightIt::weightit(
  formula = ps_formula,
  data = data2,
  method = "glm"
)

# truncate at 2.5 and 97.5 percentiles
lower <- quantile(weights_out2$weights, 0.025)
upper <- quantile(weights_out2$weights, 0.975)
weights_out2$weights <- pmin(pmax(weights_out2$weights, lower), upper)

summary(weights_out2$weights)


weights_out2$treat <- if_else(weights_out2$treat == 1, "Novel GLP1", "DPP4")

p <- bal.plot(
  weights_out2,
  var.name = "prop.score",
  which = "both",
  type = "density",
  mirror = FALSE,
  grid = FALSE,
  legend.labels = c("Novel GLP1", "DPP4"),
  position = "bottom"
)

p <- p + 
  ggtitle("Propensity Score Distributional Balance") +
  xlab("Propensity score") +
  ylab("Density")

p

ggsave(file.path(results_dir, "ps_bal2.png"), plot = p, width = 6, height = 4, units = "in", bg = "white")
dev.off()
rm(p)

bal_tab2 <- bal.tab(ps_formula,
                    data = data2,
                    weights = 'iptw',
                    binary = 'std',
                    continuous = 'std',
                    stats = c('mean.diffs'),
                    s.d.denom = 'pooled',
                    un = TRUE,
                    thresholds = c(m=0.1)
)

bal2 <- bal_tab2$Balance %>%
  dplyr::rename('SMD (Unadjusted)' = Diff.Un,
                'SMD (IPTW)' = Diff.Adj,
                'Balance' = M.Threshold)

bal2$Variable <- row.names(bal2)
write.table(bal2, file.path(results_dir, "bal2.csv"), col.names = T, row.names = T, append = F, sep = ',')

p <- love.plot(
  bal_tab2, 
  binary = "std",
  thresholds = c(m = .1),
  title = "Covariate Balance with IPTW",
  position = "bottom",
  var.order = "unadjusted",
  size = 2
)

p <- p + theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color = "black"),
  
  axis.text.x = element_text(size = 8),
  axis.text.y = element_text(size = 8),
  axis.title.x = element_text(size = 8, face = 'bold'),
  plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
  
  legend.position = "bottom", 
  legend.title = element_text(size = 8, face = 'bold'),
  legend.text = element_text(size = 8),
  legend.box.margin = margin(0, 0, 0, 0),
  legend.margin = margin(0, 0, 0, 0),
  legend.key.size = unit(0.5, "lines")
) 

p

ggsave(file.path(results_dir, "loveplot2.png"), plot = p, width = 6, height = 12, units = "in", bg = 'white')
dev.off()
rm(p)

#data3: sulfo (ref) vs novel GLP1s (comp)
data3 <- data %>%
  filter(novel_flag == 1| index_class == "SULFO") %>%
  mutate(trt = as.factor(if_else(novel_flag == 1, 1, 0)))

ps_mod_crude3 <- glm(trt ~ 1, data = data3, family = binomial)

ps_mod_adj3 <- glm(ps_formula,
                   family = binomial,
                   data = data3
)

data3$ps <- predict(ps_mod_adj3, type = "response")
data3$ps_marg <- predict(ps_mod_crude3, type = "response")

# Sensitivity: asymmetric PS trimming at trimming_pct / (1 - trimming_pct).
if (isTRUE(SCENARIO$weight_trimming)) {
  .upper_q <- 1 - SCENARIO$trimming_pct
  .lower_q <- SCENARIO$trimming_pct
  perc_upper <- data3 %>% filter(trt == 0) %>%
    summarise(perc_upper = quantile(ps, .upper_q, na.rm = TRUE))
  perc_lower <- data3 %>% filter(trt == 1) %>%
    summarise(perc_lower = quantile(ps, .lower_q, na.rm = TRUE))
  data3 %<>%
    mutate(
      perc_upper = round(perc_upper$perc_upper, 4),
      perc_lower = round(perc_lower$perc_lower, 4)
    )
  data3 <- data3 %>% filter(ps < perc_upper & ps > perc_lower)
  rm(perc_upper, perc_lower)
}

data3$iptw <- if_else(data3$trt ==1, 1/data3$ps, 1/(1-data3$ps))
data3$siptw <- data3$iptw * data3$ps_marg    

summary(data1$siptw)
test <- data1 %>%
  filter(siptw > 30)

summary(test)

summary(data3$iptw)
sd(data3$iptw)

summary(data3$siptw)
sd(data3$siptw)

# truncate at 2.5 and 97.5 percentiles
data3$iptw_notrunc <- data3$iptw
data3$siptw_notrunc <- data3$siptw

lower_bound <- quantile(data3$siptw, 0.025)
upper_bound <- quantile(data3$siptw, 0.975)
data3$siptw <- pmin(pmax(data3$siptw, lower_bound), upper_bound)
rm(lower_bound, upper_bound)

lower_bound <- quantile(data3$iptw, 0.025)
upper_bound <- quantile(data3$iptw, 0.975)
data3$iptw <- pmin(pmax(data3$iptw, lower_bound), upper_bound)

weights_out3 <- WeightIt::weightit(
  formula = ps_formula,
  data = data3,
  method = "glm"
)

# truncate at 2.5 and 97.5 percentiles
lower <- quantile(weights_out3$weights, 0.025)
upper <- quantile(weights_out3$weights, 0.975)
weights_out3$weights <- pmin(pmax(weights_out3$weights, lower), upper)

summary(weights_out3$weights)


weights_out3$treat <- if_else(weights_out3$treat == 1, "Novel GLP1", "Sulfonylureas")

p <- bal.plot(
  weights_out3,
  var.name = "prop.score",
  which = "both",
  type = "density",
  mirror = FALSE,
  grid = FALSE,
  legend.labels = c("Novel GLP1", "Sulfonylureas"),
  position = "bottom"
)

p <- p + 
  ggtitle("Propensity Score Distributional Balance") +
  xlab("Propensity score") +
  ylab("Density")

p

ggsave(file.path(results_dir, "ps_bal3.png"), plot = p, width = 6, height = 4, units = "in", bg = "white")
dev.off()

bal_tab3 <- bal.tab(ps_formula,
                    data = data3,
                    weights = 'iptw',
                    binary = 'std',
                    continuous = 'std',
                    stats = c('mean.diffs'),
                    s.d.denom = 'pooled',
                    un = TRUE,
                    thresholds = c(m=0.1)
)

bal3 <- bal_tab3$Balance %>%
  dplyr::rename('SMD (Unadjusted)' = Diff.Un,
                'SMD (IPTW)' = Diff.Adj,
                'Balance' = M.Threshold)

bal3$Variable <- row.names(bal3)
write.table(bal3, file.path(results_dir, "bal3.csv"), col.names = T, row.names = T, append = F, sep = ',')

p <- love.plot(
  bal_tab3, 
  binary = "std",
  thresholds = c(m = .1),
  title = "Covariate Balance with IPTW",
  position = "bottom",
  var.order = "unadjusted",
  size = 2
)

p <- p + theme(
  panel.border = element_blank(),
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  axis.line = element_line(color = "black"),
  
  axis.text.x = element_text(size = 8),
  axis.text.y = element_text(size = 8),
  axis.title.x = element_text(size = 8, face = 'bold'),
  plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
  
  legend.position = "bottom", 
  legend.title = element_text(size = 8, face = 'bold'),
  legend.text = element_text(size = 8),
  legend.box.margin = margin(0, 0, 0, 0),
  legend.margin = margin(0, 0, 0, 0),
  legend.key.size = unit(0.5, "lines")
) 

p

ggsave(file.path(results_dir, "loveplot3.png"), plot = p, width = 6, height = 12, units = "in", bg = 'white')
dev.off()
rm(p)

```


## COMPUTE ENTROPY BALANCING WEIGHTS 
These are not used for the main analysis but are used for one of the sensitivity analyses (ebal weights instead of IPTW). 

```{r}
#| label: compute-ebal-weights
# data 1:
ebal_weights_out1 <- WeightIt::weightit(
  formula = ps_formula,
  data = data1,
  method = "ebal"
)

summary(ebal_weights_out1$weights)

data1$ebal <- ebal_weights_out1$weights

bal_tab1 <- bal.tab(ps_formula,
                    data = data1,
                    weights = 'ebal',
                    binary = 'std',
                    continuous = 'std',
                    stats = c('mean.diffs'),
                    s.d.denom = 'pooled',
                    un = TRUE,
                    thresholds = c(m=0.1)
                    )

bal1 <- bal_tab1$Balance %>%
  dplyr::rename('SMD (Unadjusted)' = Diff.Un,
                'SMD (IPTW)' = Diff.Adj,
                'Balance' = M.Threshold)

bal1$Variable <- row.names(bal1)
write.table(bal1, file.path(results_dir, "ebal1.csv"), col.names = T, row.names = T, append = F, sep = ',')

# data 2: 
ebal_weights_out2 <- WeightIt::weightit(
  formula = ps_formula,
  data = data2,
  method = "ebal"
)

summary(ebal_weights_out2$weights)

data2$ebal <- ebal_weights_out2$weights

bal_tab2 <- bal.tab(ps_formula,
                    data = data2,
                    weights = 'ebal',
                    binary = 'std',
                    continuous = 'std',
                    stats = c('mean.diffs'),
                    s.d.denom = 'pooled',
                    un = TRUE,
                    thresholds = c(m=0.1)
                    )

bal2 <- bal_tab2$Balance %>%
  dplyr::rename('SMD (Unadjusted)' = Diff.Un,
                'SMD (IPTW)' = Diff.Adj,
                'Balance' = M.Threshold)

bal2$Variable <- row.names(bal1)
write.table(bal2, file.path(results_dir, "ebal2.csv"), col.names = T, row.names = T, append = F, sep = ',')

# data 3: 
ebal_weights_out3 <- WeightIt::weightit(
  formula = ps_formula,
  data = data3,
  method = "ebal"
)

summary(ebal_weights_out3$weights)

data3$ebal <- ebal_weights_out3$weights

bal_tab3 <- bal.tab(ps_formula,
                    data = data3,
                    weights = 'ebal',
                    binary = 'std',
                    continuous = 'std',
                    stats = c('mean.diffs'),
                    s.d.denom = 'pooled',
                    un = TRUE,
                    thresholds = c(m=0.1)
                    )

bal3 <- bal_tab3$Balance %>%
  dplyr::rename('SMD (Unadjusted)' = Diff.Un,
                'SMD (IPTW)' = Diff.Adj,
                'Balance' = M.Threshold)

bal3$Variable <- row.names(bal3)
write.table(bal3, file.path(results_dir, "ebal3.csv"), col.names = T, row.names = T, append = F, sep = ',')
```

## SAVE COHORT DATA WITH WEIGHTS

```{r}
#| label: combine-and-save-weights
data1 %<>%
  mutate(
    wfinal = ipcw_switch * ipcw_disc * iptw,
    swfinal = ipcw_switch * ipcw_disc * siptw,
    ebalfinal = ipcw_switch + ipcw_disc * ebal 
  )


data2 %<>%
  mutate(
    wfinal = ipcw_switch * ipcw_disc * iptw,
    swfinal = ipcw_switch * ipcw_disc * siptw,
    ebalfinal = ipcw_switch + ipcw_disc * ebal
  )


data3 %<>%
  mutate(
    wfinal = ipcw_switch * ipcw_disc * iptw,
    swfinal = ipcw_switch * ipcw_disc * siptw,
    ebalfinal = ipcw_switch + ipcw_disc * ebal
  )

# save all cohort dataframes
# ready for description + analysis
saveRDS(data, file.path(output_dir, "data_novel.rds")) # all patients
saveRDS(data1, file.path(output_dir, "data1_novel.rds")) # novel GLP1s vs sglt2
saveRDS(data2, file.path(output_dir, "data2_novel.rds")) # novel GLP1s vs dpp4
saveRDS(data3, file.path(output_dir, "data3_novel.rds")) # novel GLP1s vs sulfo
```


## SAVE MODELS & DESCRIBE WEIGHTS

```{r}
#| label: describe-weights
# save models

sink(file.path(results_dir, "iptw_models_novel.txt"))

print(ps_formula)

cat("=" , rep("=", 70), "\n", sep = "")
cat("IPTW MODELS\n")
cat("Generated on:", format(Sys.time(), "%Y-%m-%d %H:%M:%S"), "\n")
cat("=" , rep("=", 70), "\n\n", sep = "")


cat("\n\n")
cat("MODEL 1\n")
cat("-" , rep("-", 70), "\n", sep = "")

print(summary(ps_mod_adj1))

cat("\n\n")
cat("MODEL 2\n")
cat("-" , rep("-", 70), "\n", sep = "")

print(summary(ps_mod_adj2))

cat("\n\n")
cat("MODEL 3\n")
cat("-" , rep("-", 70), "\n", sep = "")

print(summary(ps_mod_adj3))

sink()



#### DESCRIBE THE WEIGHTS ####

weights_summary_d1 <- data1 %>%
  pivot_longer(cols = c(iptw, siptw, wfinal, swfinal, ebal, ebalfinal),
               names_to = "Weight_Type", values_to = "value") %>%
  group_by(Weight_Type) %>%
  summarise(
    Mean = round(mean(value, na.rm = TRUE), 3),
    Median = round(median(value, na.rm = TRUE), 3),
    SD = round(sd(value, na.rm = TRUE), 1),
    Min = round(min(value, na.rm = TRUE), 3),
    Max = round(max(value, na.rm = TRUE), 3),
    Q1 = round(quantile(value, 0.25, na.rm = TRUE), 3),
    Q3 = round(quantile(value, 0.75, na.rm = TRUE), 3),
    IQR = round(IQR(value, na.rm = TRUE), 3)
  ) %>%
  mutate(Group = "Semaglutide vs SGLT2")


weights_summary_d2 <- data2 %>%
  pivot_longer(cols = c(iptw, siptw, wfinal, swfinal, ebal, ebalfinal),
               names_to = "Weight_Type", values_to = "value") %>%
  group_by(Weight_Type) %>%
  summarise(
    Mean = round(mean(value, na.rm = TRUE), 3),
    Median = round(median(value, na.rm = TRUE), 3),
    SD = round(sd(value, na.rm = TRUE), 1),
    Min = round(min(value, na.rm = TRUE), 3),
    Max = round(max(value, na.rm = TRUE), 3),
    Q1 = round(quantile(value, 0.25, na.rm = TRUE), 3),
    Q3 = round(quantile(value, 0.75, na.rm = TRUE), 3),
    IQR = round(IQR(value, na.rm = TRUE), 3)
  ) %>%
  mutate(Group = "Semaglutide vs DPP4")


weights_summary_d3 <- data3 %>%
  pivot_longer(cols = c(iptw, siptw, wfinal, swfinal, ebal, ebalfinal),
               names_to = "Weight_Type", values_to = "value") %>%
  group_by(Weight_Type) %>%
  summarise(
    Mean = round(mean(value, na.rm = TRUE), 3),
    Median = round(median(value, na.rm = TRUE), 3),
    SD = round(sd(value, na.rm = TRUE), 1),
    Min = round(min(value, na.rm = TRUE), 3),
    Max = round(max(value, na.rm = TRUE), 3),
    Q1 = round(quantile(value, 0.25, na.rm = TRUE), 3),
    Q3 = round(quantile(value, 0.75, na.rm = TRUE), 3),
    IQR = round(IQR(value, na.rm = TRUE), 3)
  ) %>%
  mutate(Group = "Semaglutide vs Sulfos")


write.csv(weights_summary_d1, file.path(results_dir, "iptw_and_final_weights_d1.csv"), row.names = FALSE)
write.csv(weights_summary_d2, file.path(results_dir, "iptw_and_final_weights_d2.csv"), row.names = FALSE)
write.csv(weights_summary_d3, file.path(results_dir, "iptw_and_final_weights_d3.csv"), row.names = FALSE)
```