---
title: "Survival Analyses - Novel GLP1s vs other classes"
author: "Gwen Aubrac"
format: html
embed-resources: true
---

## Overview

Perform ITT survival analyses using Cox proportional hazards model with IPTW weights:
- For multiple outcomes: dementia outcomes and subtypes (all-cause, lewy body, Alzheimer), negative control outcomes, and sentinel measures of dementia (order for neuroimaging, Parkinson's disease, and MCI). For sentinel measures, patients are additionally censored when they experience all-cause dementia. 
- In ITT, patients are censored at coverage end, study end, follow-up cap (see below), date of death, or occurrence of the outcome.
- Follow-up starts after 90 days to reduce the influence of early events. Patients with an early event during this window were excluded from the analysis. A negative follow-up means that the patient left the cohort (due to death, end of coverage, end of study, cap of follow-up, occurrence of the event) during that period of first 90 days. Those patients were excluded and did not contribute to the numerator nor denominator.

Perform AT survival analyses using Cox proportional hazards model with IPCW weights: 
- Patients are censored at ITT date or additionally when they switch or discontinue treatment. 
- Follow-up starts after 90 days to reduce the influence of early events. 

Follow-up is capped at 2 years (730 days) and this includes the 90 days of follow-up excluded at the start.

Notes on coding: The analyses are done using a loop to analyze all outcomes for all contrasts in one go. The "cohort names" are:
- data1: novel GLP1 (sema + tirze) vs SGLT2
- data2: novel GLP1 vs DPP4
- data3: nvel GLP1 vs sulfo

**Sensitivity-controlled options** (set by `SCENARIO` - see `sensitivity.R`):
- `weight_method` (`"iptw"` or `"ebal"` or `"ow"`): which weight column is used for the weighted Cox models. `"iptw"` uses `siptw` for ITT and (depending on `use_ipcw`) `siptw` or `swfinal` for PP. `"ebal"` uses `ebal` for ITT and (depending on `use_ipcw`) `ebal` or `ebalfinal` for PP. `"ow"` uses `ow` for ITT and (depending on `use_ipcw`) `ow` or `owfinal` for PP.
- `use_ipcw`: if TRUE, the PP analysis combines IPTW (or entropy) with IPCW for switch/discontinuation.
- `followup_start` (`"after_grace"` or `"index_date"`): start follow-up `index + 90d` (default) or at `index`.
- `extra_followup_days` (integer): when 0, keep patients with any positive follow-up (original behavior); when N>0, require `follow_up >= N` days in both ITT and PP cohorts.

```{r}
#| label: survival-models
#### LOAD DATA ####

project_root <- Sys.getenv("PROJECT_ROOT")
if (nchar(project_root) == 0) stop("PROJECT_ROOT is not set. To run this file directly in an R session, source code/00-setup.R first in the same session (it loads .Renviron). Otherwise launch the full pipeline via run_all.R / run_all.command / run_all.bat.")
output_dir      <- Sys.getenv("OUTPUT_DIR",      unset = file.path(project_root, "output"))
main_output_dir <- Sys.getenv("MAIN_OUTPUT_DIR", unset = output_dir)
results_dir <- Sys.getenv("RESULTS_DIR", unset = file.path(project_root, "results", "main"))
is_main <- !nzchar(Sys.getenv("SCENARIO_NAME")) ||
  identical(Sys.getenv("SCENARIO_NAME"), "main")
out_path <- function(name) {
  p <- normalizePath(file.path(output_dir, name), winslash = "/", mustWork = FALSE)
  if (file.exists(p)) return(p)
  normalizePath(file.path(main_output_dir, name), winslash = "/", mustWork = FALSE)
}

source(file.path(project_root, "sensitivity.R"))
SCENARIO <- get_scenario()

# Resolve weight columns from scenario settings. `siptw`/`swfinal`/`ebal`/`ebalfinal`
# are columns written by 16-compute-iptw.qmd; see there for their construction.
itt_weight_col <- switch(SCENARIO$weight_method,
  "ebal" = "ebal",
  "ow" = "ow",
  "siptw" # default
)

pp_weight_col <- switch(SCENARIO$weight_method,
  "ebal" = if (isTRUE(SCENARIO$use_ipcw)) "ebalfinal" else "ebal",
  "ow" = if (isTRUE(SCENARIO$use_ipcw)) "owfinal"   else "ow",
  if (isTRUE(SCENARIO$use_ipcw)) "swfinal" else "siptw" # default
)

# Follow-up start offset in days from index.
fu_offset_days <- if (identical(SCENARIO$followup_start, "index_date")) 0L else 90L

library(dplyr)
library(magrittr)
library(table1)
library(lubridate)
library(survival)
library(survminer)
library(tidyr)
library(broom)
library(openxlsx)

end_date <- ymd(Sys.getenv("STUDY_END", unset = "2025-12-31"))

data     <- readRDS(out_path("data_novel.rds"))
data1    <- readRDS(out_path("data1_novel.rds"))
data2    <- readRDS(out_path("data2_novel.rds"))
data3    <- readRDS(out_path("data3_novel.rds"))
cov_list <- readRDS(out_path("cov_list_novel.rds"))

max_fu <- 730 # truncate follow-up up to 2 years after entry

#### SURVIVAL ANALYSES ####

# here, patients will not be censored if they experience all-cause dementia
# since all-cause dementia or its subtypes are the events of interest

# we are also incorporating the negative control outcomes in this loop
# since similarly, we would not censor at all-cause dementia for those

# save results from all models
all_cox_results <- list()

# save follow-up times and event count for each analysis
followup_times <- data.frame(
  cohort = NA,
  approach = NA,
  outcome = NA,
  mean = NA,
  min = NA,
  max = NA,
  median = NA,
  q1 = NA,
  q3 = NA,
  n_trt = NA,
  n_comp = NA,
  n_events = NA,
  n_events_trt = NA,
  n_events_comp = NA
)

# save proportional hazards assumption test for each model
cox_ph <- data.frame(
  cohort = NA,
  approach = NA,
  outcome = NA,
  pvalue = NA
)

# define outcomes as named within dataset var names:
# ex: "dementia_allcause" with var names "outcome_dementia_allcause_date" and res will be saved as "dementia_allcause"

dementia_outcomes <- c(
  # dementia outcomes:
  "dementia_allcause",

  # negative control outcomes:
  "appendicitis",
  "tooth",
  "bell",
  "basal"
)

for (outcome in c(dementia_outcomes)) {
  
  for (cohort_name in c("data1", "data2", "data3")) {

    cohort <- get(cohort_name)
    
    date_name <- paste0("outcome_", outcome, "_date")
    
    
    if (cohort_name == "data1") {
      ref <- "SGLT2"
      index <- "Semaglutide/Tirzepatide"
    } else if (cohort_name == "data2") {
      ref <- "DPP4"
      index <- "Semaglutide/Tirzepatide"
    } else if (cohort_name == "data3") {
      ref <- "Sulfonylureas"
      index <- "Semaglutide/Tirzepatide"
    }

    
    ### ITT APPROACH ###
    
    # describe follow-up (ITT)
    cohort$study_end <- end_date
    
    cohort$itt_exit_date <- pmin(
      cohort$dod, # date of death
      cohort$cov_end, # end of coverage
      cohort$study_end, # end of study
      cohort$index_date + max_fu, 
      na.rm = TRUE
    )
    
    summary(cohort$itt_exit_date)

    # Start follow-up `fu_offset_days` after index (90 for main, 0 for
    # the sens7_index_followup scenario), to reduce the influence of early
    # events likely driven by confounding by indication / reverse causality.
    cohort$f_up_start <- cohort$index_date + fu_offset_days

    # create indicator for itt_event
    # no event if no date of event or if occurred after ITT exit
    # !!sym(date_name) takes the date for the event being analyzed in this loop

    cohort %<>%
      mutate (itt_event = if_else(!is.na(!!sym(date_name)) & !!sym(date_name) < itt_exit_date & !!sym(date_name) >= f_up_start, 1, 0),
              itt_event_date = if_else(itt_event == 1, !!sym(date_name), NA))
    
    table(cohort$itt_event)
    
    # update itt exit date if there was an event
    # so that follow-up additionally ends at event occurrence
    cohort %<>%
      mutate(itt_exit_date = if_else(itt_event == 1, itt_event_date, itt_exit_date))
    
    cohort$itt_follow_up <- time_length(interval(as.Date(cohort$f_up_start), as.Date(cohort$itt_exit_date)), 'days')
    summary(cohort$itt_follow_up)

    # extra_followup_days == 0 keeps the original >0 filter exactly;
    # extra_followup_days > 0 requires >= N days (sens8_6mo_followup).
    if (SCENARIO$extra_followup_days <= 0) {
      cohort %<>% filter(itt_follow_up > 0)
    } else {
      cohort %<>% filter(itt_follow_up >= SCENARIO$extra_followup_days)
    }

    # fill in info on follow-up
    followup_times <- rbind(
      followup_times,
      data.frame(
        cohort = cohort_name,
        approach = "itt",
        outcome = outcome,
        mean = mean(cohort$itt_follow_up),
        min = min(cohort$itt_follow_up),
        max = max(cohort$itt_follow_up),
        median = median(cohort$itt_follow_up),
        q1 = quantile(cohort$itt_follow_up, 0.25),
        q3 = quantile(cohort$itt_follow_up, 0.75),
        n_trt = sum(cohort$trt==1),
        n_comp = sum(cohort$trt==0),
        n_events = sum(cohort$itt_event),
        n_events_trt = sum(cohort$itt_event[cohort$trt==1]),
        n_events_comp = sum(cohort$itt_event[cohort$trt==0])
      )
    )
    
    
    # fit survival models
    # crude ITT model: 
    cox_mod_itt_crude <- coxph(Surv(itt_follow_up, itt_event) ~ trt, 
                         data = cohort, 
                         cluster = PATIENT_ID, 
                         robust = TRUE)

    coxph_p <- cox.zph(cox_mod_itt_crude)$table["trt", "p"]
    
    cox_ph <- rbind(
      cox_ph,
      data.frame(
        cohort = cohort_name,
        approach = "itt crude",
        outcome = outcome,
        pvalue = coxph_p
      )
    )
    
    cox_mod_itt_res_crude <- tidy(cox_mod_itt_crude, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "itt_crude"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results[[paste(outcome, cohort_name, "itt_crude", sep = "_")]] <- cox_mod_itt_res_crude
    
    km_curve_itt_crude <- survfit(Surv(itt_follow_up, itt_event) ~ trt, 
                            data = cohort)
    
    
    
    
    # Weighted ITT model; weight column chosen from SCENARIO$weight_method
    # (siptw for "iptw" scenarios, ebal for "ebal" scenarios, ow for "ow" scenarios).
    cox_mod_itt <- coxph(Surv(itt_follow_up, itt_event) ~ trt,
                         data = cohort,
                         weights = cohort[[itt_weight_col]],
                         cluster = PATIENT_ID,
                         robust = TRUE)

    coxph_p <- cox.zph(cox_mod_itt)$table["trt", "p"]
    
    cox_ph <- rbind(
      cox_ph,
      data.frame(
        cohort = cohort_name,
        approach = "itt IPTW",
        outcome = outcome,
        pvalue = coxph_p
      )
    )

    cox_mod_itt_res <- tidy(cox_mod_itt, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "itt"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results[[paste(outcome, cohort_name, "itt", sep = "_")]] <- cox_mod_itt_res
    
    
    
    
    
    ### PER PROTOCOL APPROACH ###
    
    # describe follow-up (PP)
    # same as above, but additionally censoring at switch and disc
    cohort$pp_exit_date <- pmin(
      cohort$dod,
      cohort$cov_end,
      cohort$study_end,
      cohort$switch_date,
      cohort$disc_date,
      cohort$index_date + max_fu,
      na.rm = TRUE
    )
    
    summary(cohort$pp_exit_date)
    
    # create indicator for pp_event
    # no event if no date of event or if occurred after PP exit
    cohort %<>%
      mutate (pp_event = if_else(!is.na(!!sym(date_name)) & !!sym(date_name) < pp_exit_date & !!sym(date_name) >= f_up_start, 1, 0),
              pp_event_date = if_else(pp_event == 1, !!sym(date_name), NA))
    
    table(cohort$pp_event)
    
    # update pp exit date if there was an event
    # so that follow-up additionally ends at event occurrence
    cohort %<>%
      mutate(pp_exit_date = if_else(pp_event == 1, pp_event_date, pp_exit_date))
    
    
    cohort$pp_follow_up <- time_length(interval(as.Date(cohort$f_up_start), as.Date(cohort$pp_exit_date)), 'days')
    summary(cohort$pp_follow_up)

    if (SCENARIO$extra_followup_days <= 0) {
      cohort %<>% filter(pp_follow_up > 0)
    } else {
      cohort %<>% filter(pp_follow_up >= SCENARIO$extra_followup_days)
    }
  
    
    # fill in info on follow-up
    followup_times <- rbind(
      followup_times,
      data.frame(
        cohort = cohort_name,
        approach = "pp",
        outcome = outcome,
        mean = mean(cohort$pp_follow_up),
        min = min(cohort$pp_follow_up),
        max = max(cohort$pp_follow_up),
        median = median(cohort$pp_follow_up),
        q1 = quantile(cohort$pp_follow_up, 0.25),
        q3 = quantile(cohort$pp_follow_up, 0.75),
        n_trt = sum(cohort$trt==1),
        n_comp = sum(cohort$trt==0),
        n_events = sum(cohort$pp_event),
        n_events_trt = sum(cohort$pp_event[cohort$trt==1]),
        n_events_comp = sum(cohort$pp_event[cohort$trt==0])
      )
    )
    
    
    # fit survival models
    # crude model:
    cox_mod_pp_crude <- coxph(Surv(pp_follow_up, pp_event) ~ trt, 
                        data = cohort,
                        cluster = PATIENT_ID, 
                        robust = TRUE)

    coxph_p <- cox.zph(cox_mod_pp_crude)$table["trt", "p"]
    
    cox_ph <- rbind(
      cox_ph,
      data.frame(
        cohort = cohort_name,
        approach = "pp crude",
        outcome = outcome,
        pvalue = coxph_p
      )
    )
    
    cox_mod_pp_res_crude <- tidy(cox_mod_pp_crude, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "pp_crude"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results[[paste(outcome, cohort_name, "pp_crude", sep = "_")]] <- cox_mod_pp_res_crude
    

    
    # Weighted PP model; weight column depends on SCENARIO$weight_method and
    # SCENARIO$use_ipcw:
    #   iptw, no ipcw: uses siptw (main)
    #   iptw, ipcw: uses swfinal (sens4_pp_ipcw)
    #   ebal, no ipcw: uses ebal (sens1_ebal)
    #   ebal, ipcw: uses ebalfinal
    #   ow, no ipcw: uses ow (sens1_ow)
    #   ow, ipcw: uses owfinal
    cox_mod_pp <- coxph(Surv(pp_follow_up, pp_event) ~ trt,
                            data = cohort,
                            weights = cohort[[pp_weight_col]],
                            cluster = PATIENT_ID,
                            robust = TRUE)

    coxph_p <- cox.zph(cox_mod_pp)$table["trt", "p"]
    
    cox_ph <- rbind(
      cox_ph,
      data.frame(
        cohort = cohort_name,
        approach = "pp IPTW+IPCW",
        outcome = outcome,
        pvalue = coxph_p
      )
    )
    
    cox_mod_pp_res <- tidy(cox_mod_pp, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "pp"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results[[paste(outcome, cohort_name, "pp", sep = "_")]] <- cox_mod_pp_res
    
    
  }
}


rm(cohort, cohort_name, outcome)
gc()


#### SAVE MODEL RESULTS ####

write.csv(
  followup_times,
  file.path(results_dir, "followup_times.csv"),
  row.names = FALSE
)

cox_ph %<>%
  mutate(
    flag = if_else(pvalue<0.05, 1, 0)
  )

write.csv(
  cox_ph,
  file.path(results_dir, "cox_ph.csv"),
  row.names = FALSE
)

summary_table <- bind_rows(all_cox_results)

summary_table <- summary_table %>%
  mutate(across(where(is.numeric), ~round(., 2)))


write.csv(
  summary_table,
  file.path(results_dir, "all_cox_results.csv"),
  row.names = FALSE
)

saveRDS(
  summary_table,
  file.path(results_dir, "all_cox_results_novel.rds")
)

```

## FALSIFICATION TEST: RISK IN FIRST 90 DAYS

As a falsification test, we can look at the risk of dementia in the first 90 days of follow-up, when a true preventive effect is implausible. If a treatment effect is found, it would suggest confounding as the groups have a different baseline risk of dementia. 

```{r}
# save results from all models
all_cox_results_fals <- list()

# save follow-up times and event count for each analysis
followup_times_fals <- data.frame(
  cohort = NA,
  approach = NA,
  outcome = NA,
  mean = NA,
  min = NA,
  max = NA,
  median = NA,
  q1 = NA,
  q3 = NA,
  n_trt = NA,
  n_comp = NA,
  n_events = NA,
  n_events_trt = NA,
  n_events_comp = NA
)

# save proportional hazards assumption test for each model
cox_ph_fals <- data.frame(
  cohort = NA,
  approach = NA,
  outcome = NA,
  pvalue = NA
)

# define outcomes as named within dataset var names:
# ex: "dementia_allcause" with var names "outcome_dementia_allcause_date" and res will be saved as "dementia_allcause"

dementia_outcomes <- c(
  # dementia outcomes:
  "dementia_allcause",
)

for (outcome in c(dementia_outcomes)) {
  
  for (cohort_name in c("data1", "data2", "data3")) {

    cohort <- get(cohort_name)
    
    date_name <- paste0("outcome_", outcome, "_date")
    
    
    if (cohort_name == "data1") {
      ref <- "SGLT2"
      index <- "Semaglutide/Tirzepatide"
    } else if (cohort_name == "data2") {
      ref <- "DPP4"
      index <- "Semaglutide/Tirzepatide"
    } else if (cohort_name == "data3") {
      ref <- "Sulfonylureas"
      index <- "Semaglutide/Tirzepatide"
    }

    
    ### ITT APPROACH ###
    
    # describe follow-up (ITT)
    cohort$study_end <- end_date
    
    cohort$itt_exit_date <- pmin(
      cohort$dod, # date of death
      cohort$cov_end, # end of coverage
      cohort$study_end, # end of study
      cohort$index_date + 90, # at most 90 days after follow-up
      na.rm = TRUE
    )
    
    summary(cohort$itt_exit_date)

    # Start follow-up `fu_offset_days` on index date
    cohort$f_up_start <- cohort$index_date 

    # create indicator for itt_event
    # no event if no date of event or if occurred after ITT exit
    # !!sym(date_name) takes the date for the event being analyzed in this loop

    cohort %<>%
      mutate (itt_event = if_else(!is.na(!!sym(date_name)) & !!sym(date_name) < itt_exit_date & !!sym(date_name) >= f_up_start, 1, 0),
              itt_event_date = if_else(itt_event == 1, !!sym(date_name), NA))
    
    table(cohort$itt_event)
    
    # update itt exit date if there was an event
    # so that follow-up additionally ends at event occurrence
    cohort %<>%
      mutate(itt_exit_date = if_else(itt_event == 1, itt_event_date, itt_exit_date))
    
    cohort$itt_follow_up <- time_length(interval(as.Date(cohort$f_up_start), as.Date(cohort$itt_exit_date)), 'days')
    summary(cohort$itt_follow_up)

    cohort %<>% filter(itt_follow_up > 0)

    # fill in info on follow-up
    followup_times_fals <- rbind(
      followup_times_fals,
      data.frame(
        cohort = cohort_name,
        approach = "itt",
        outcome = outcome,
        mean = mean(cohort$itt_follow_up),
        min = min(cohort$itt_follow_up),
        max = max(cohort$itt_follow_up),
        median = median(cohort$itt_follow_up),
        q1 = quantile(cohort$itt_follow_up, 0.25),
        q3 = quantile(cohort$itt_follow_up, 0.75),
        n_trt = sum(cohort$trt==1),
        n_comp = sum(cohort$trt==0),
        n_events = sum(cohort$itt_event),
        n_events_trt = sum(cohort$itt_event[cohort$trt==1]),
        n_events_comp = sum(cohort$itt_event[cohort$trt==0])
      )
    )
    
    
    # fit survival models
    # crude ITT model: 
    cox_mod_itt_crude <- coxph(Surv(itt_follow_up, itt_event) ~ trt, 
                         data = cohort, 
                         cluster = PATIENT_ID, 
                         robust = TRUE)

    coxph_p <- cox.zph(cox_mod_itt_crude)$table["trt", "p"]
    
    cox_ph_fals <- rbind(
      cox_ph_fals,
      data.frame(
        cohort = cohort_name,
        approach = "itt crude",
        outcome = outcome,
        pvalue = coxph_p
      )
    )
    
    cox_mod_itt_res_crude <- tidy(cox_mod_itt_crude, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "itt_crude"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results_fals[[paste(outcome, cohort_name, "itt_crude", sep = "_")]] <- cox_mod_itt_res_crude
    
    km_curve_itt_crude <- survfit(Surv(itt_follow_up, itt_event) ~ trt, 
                            data = cohort)
    
    
    
    
    # Weighted ITT model
    cox_mod_itt <- coxph(Surv(itt_follow_up, itt_event) ~ trt,
                         data = cohort,
                         weights = "siptw",
                         cluster = PATIENT_ID,
                         robust = TRUE)

    coxph_p <- cox.zph(cox_mod_itt)$table["trt", "p"]
    
    cox_ph_fals <- rbind(
      cox_ph_fals,
      data.frame(
        cohort = cohort_name,
        approach = "itt IPTW",
        outcome = outcome,
        pvalue = coxph_p
      )
    )

    cox_mod_itt_res <- tidy(cox_mod_itt, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "itt"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results_fals[[paste(outcome, cohort_name, "itt", sep = "_")]] <- cox_mod_itt_res
  }
}


rm(cohort, cohort_name, outcome)
gc()


#### SAVE MODEL RESULTS ####

write.csv(
  followup_times_fals,
  file.path(results_dir, "followup_times_fals.csv"),
  row.names = FALSE
)

cox_ph_fals %<>%
  mutate(
    flag = if_else(pvalue<0.05, 1, 0)
  )

write.csv(
  cox_ph_fals,
  file.path(results_dir, "cox_ph_fals.csv"),
  row.names = FALSE
)

summary_table_fals <- bind_rows(all_cox_results_fals)

summary_table_fals <- summary_table_fals %>%
  mutate(across(where(is.numeric), ~round(., 2)))


write.csv(
  summary_table_fals,
  file.path(results_dir, "all_cox_results_fals.csv"),
  row.names = FALSE
)

saveRDS(
  summary_table_fals,
  file.path(results_dir, "all_cox_results_novel_fals.rds")
)
```

## FINAL TABLE FOR MANUSCRIPT 

```{r}
#| label: save-final-table2
#### SAVE FINAL TABLE 2 ####

final_tab2_data1 <- summary_table %>%
  filter(cohort == "data1") %>% # keep only results for GLP1 vs SGLT2
  filter(outcome == "dementia_allcause") %>%
  filter(approach == "itt" | approach == "pp") %>% # keep only adjusted estimates
  mutate(estimate = paste0(hr, " (", ci_lower, ", ", ci_upper, ")")) %>% # format estimate and CI into same cell
  select(approach, estimate)

final_tab2_data2 <- summary_table %>%
  filter(cohort == "data2") %>% # keep only results for GLP1 vs DPP4
  filter(outcome == "dementia_allcause") %>%
  filter(approach == "itt" | approach == "pp") %>% # keep only adjusted estimates
  mutate(estimate = paste0(hr, " (", ci_lower, ", ", ci_upper, ")")) %>% # format estimate and CI into same cell
  select(approach, estimate)

final_tab2_data3 <- summary_table %>%
  filter(cohort == "data3") %>% # keep only results for GLP1 vs sulfo
  filter(outcome == "dementia_allcause") %>%
  filter(approach == "itt" | approach == "pp") %>% # keep only adjusted estimates
  mutate(estimate = paste0(hr, " (", ci_lower, ", ", ci_upper, ")")) %>% # format estimate and CI into same cell
  select(approach, estimate)

final_tab2 <- data.frame(
  glp1_vs_sglt2 = c(final_tab2_data1$estimate[final_tab2_data1$approach=="itt"], final_tab2_data1$estimate[final_tab2_data1$approach=="pp"]),
  glp1_vs_dpp4 = c(final_tab2_data2$estimate[final_tab2_data2$approach=="itt"], final_tab2_data2$estimate[final_tab2_data2$approach=="pp"]),
  glp1_vs_sulfo = c(final_tab2_data3$estimate[final_tab2_data3$approach=="itt"], final_tab2_data3$estimate[final_tab2_data3$approach=="pp"])
)

row.names(final_tab2) <- c("ITT", "PP")

write.csv(
  final_tab2,
  file.path(results_dir, "final_tab2.csv"),
  row.names = FALSE
)
```

## MANUSCRIPT TABLES

```{r}
#| label: manuscript-nco-and-results
# Manuscript tables are written only on the main run; sensitivity scenarios
# would otherwise overwrite main's manuscript copies in the shared folder.
if (!is_main) knitr::knit_exit()

manuscript_tab_dir <- file.path(project_root, "manuscript", "tables")
dir.create(manuscript_tab_dir, recursive = TRUE, showWarnings = FALSE)

write_manuscript_xlsx <- function(df, path, sheet = "Sheet1", label_col = 1) {
  val_cols <- setdiff(seq_len(ncol(df)), label_col)
  lab <- as.character(df[[label_col]]); lab[is.na(lab)] <- ""
  indented <- grepl("^[[:space:]]", lab)
  trimmed  <- trimws(lab)
  vals_empty <- vapply(seq_len(nrow(df)), function(i)
    all(is.na(df[i, val_cols]) | trimws(as.character(df[i, val_cols])) == ""),
    logical(1))
  is_section <- !indented & nzchar(trimmed) & vals_empty

  out <- df
  out[[label_col]] <- trimmed

  wb <- createWorkbook()
  addWorksheet(wb, sheet)
  writeData(wb, sheet, out, colNames = TRUE)
  nr <- nrow(out); nc <- ncol(out)

  font <- list(fontName = "Times New Roman", fontSize = 12)
  addStyle(wb, sheet, do.call(createStyle, c(font, list(valign = "center", wrapText = TRUE))),
           rows = 1:(nr + 1), cols = 1:nc, gridExpand = TRUE)
  addStyle(wb, sheet, do.call(createStyle, c(font, list(textDecoration = "bold", wrapText = TRUE))),
           rows = 1, cols = 1:nc, gridExpand = TRUE)
  for (i in seq_len(nr)) {
    st <- do.call(createStyle, c(font, list(
      halign = "left",
      indent = if (indented[i]) 1 else 0,
      textDecoration = if (is_section[i]) "bold" else NULL
    )))
    addStyle(wb, sheet, st, rows = i + 1, cols = label_col, gridExpand = TRUE)
  }
  setColWidths(wb, sheet, cols = label_col, widths = 34)
  setColWidths(wb, sheet, cols = val_cols, widths = 22)
  saveWorkbook(wb, path, overwrite = TRUE)
}

cohort_cols <- c(
  data1 = "HR (95% CI)\nGLP1s vs SGLT2s",
  data2 = "HR (95% CI)\nGLP1s vs DPP4s",
  data3 = "HR (95% CI)\nGLP1s vs Sulfonylureas"
)

fmt_hr <- function(st, app, out = "dementia_allcause") {
  vapply(names(cohort_cols), function(coh) {
    r <- st %>% filter(cohort == coh, outcome == out, approach == app)
    if (nrow(r) == 1) sprintf("%.2f (%.2f-%.2f)", r$hr, r$ci_lower, r$ci_upper)
    else NA_character_
  }, character(1))
}

#### e-Table 4: Negative Control Outcomes (IPTW-weighted ITT) ####
nco_labels <- c(
  appendicitis = "Appendicitis",
  tooth        = "Traumatic tooth fracture",
  bell         = "Bell's Palsy",
  basal        = "Basal cell carcinoma"
)

e_tab5 <- do.call(rbind, lapply(names(nco_labels), function(out) {
  c(`Negative Control Outcome` = nco_labels[[out]], fmt_hr(summary_table, "itt", out = out))
}))
e_tab5 <- as.data.frame(e_tab5, stringsAsFactors = FALSE)
colnames(e_tab5) <- c("Negative Control Outcome", unname(cohort_cols))
write_manuscript_xlsx(e_tab5,
                      file.path(manuscript_tab_dir, "e-table-5-nco.xlsx"),
                      sheet = "e-Table 5")
```

```{r}
#| label: manuscript-table-2-results
#| warning: false

# Table 2 aggregates the main analysis with every sensitivity scenario run
# so far and is only built when running main analysis
scenario_name <- Sys.getenv("SCENARIO_NAME", unset = "main")
is_main <- !nzchar(scenario_name) || identical(scenario_name, "main")

if (is_main) {
  source(file.path(project_root, "sensitivity.R"))

  rows <- list()
  rows[[length(rows) + 1]] <- c("Primary Analysis", "", "", "")
  rows[[length(rows) + 1]] <- c("  Crude", fmt_hr(summary_table, "itt_crude"))
  rows[[length(rows) + 1]] <- c("  IPTW",  fmt_hr(summary_table, "itt"))
  rows[[length(rows) + 1]] <- c("Sensitivity analyses", "", "", "")

  # Order of sensitivity rows in manuscript Table 2. 
  sens_order <- c(
    "sens1_ebal",
    "sens2_trim",
    "pp_iptw_only",
    "sens4_pp_ipcw",
    "sens5_metformin",
    "sens6_6mo_coverage",
    "sens7_index_followup",
    "sens8_6mo_followup",
    "sens9_age_cap",
    "sens10_ow",
  )
  sens_order <- c(sens_order,
                  setdiff(names(SENSITIVITY_SCENARIOS), sens_order))

  # Sensitivity scenarios save their rds to results/<scenario>/
  scenarios_root <- dirname(results_dir)
  for (nm in sens_order) {
    if (identical(nm, "pp_iptw_only")) {
      rows[[length(rows) + 1]] <- c("  Per-protocol with IPTW only",
                                    fmt_hr(summary_table, "pp"))
      next
    }
    sens_rds <- file.path(scenarios_root, nm, "all_cox_results_novel.rds")
    if (!file.exists(sens_rds)) next
    sens_st <- readRDS(sens_rds)
    app <- if (identical(nm, "sens4_pp_ipcw")) "pp" else "itt"
    rows[[length(rows) + 1]] <- c(
      paste0("  ", SENSITIVITY_SCENARIOS[[nm]]$label),
      fmt_hr(sens_st, app)
    )
  }

  tab2 <- as.data.frame(do.call(rbind, rows), stringsAsFactors = FALSE)
  colnames(tab2) <- c("Analysis", unname(cohort_cols))
  write_manuscript_xlsx(tab2,
                        file.path(manuscript_tab_dir, "table-2-results.xlsx"),
                        sheet = "Table 2")
} else {
  message("Skipping table-2-results: SCENARIO_NAME='", scenario_name,
          "' - table 2 is only assembled on the main run.")
}
```
