---
title: "Survival Analyses - Novel GLP1s vs other classes"
author: "Gwen Aubrac"
format: html
---

## Overview

Perform ITT survival analyses using Cox proportional hazards model with IPTW weights:
- For multiple outcomes: dementia outcomes and subtypes (all-cause, lewy body, Alzheimer), negative control outcomes, and sentinel measures of dementia (order for neuroimaging, Parkinson's disease, and MCI). For sentinel measures, patients are additionally censored when they experience all-cause dementia. 
- In ITT, patients are censored at coverage end, study end, follow-up cap (see below), date of death, or occurrence of the outcome.
- Follow-up starts after 90 days to reduce the influence of early events. Patients with an early event during this window were excluded from the analysis. A negative follow-up means that the patient left the cohort (due to death, end of coverage, end of study, cap of follow-up, occurrence of the event) during that period of first 90 days. Those patients were excluded and did not contribute to the numerator nor denominator.

Perform AT survival analyses using Cox proportional hazards model with IPCW weights: 
- Patients are censored at ITT date or additionally when they switch or discontinue treatment. 
- Follow-up starts after 90 days to reduce the influence of early events. 

Follow-up is capped at 2 years (730 days) and this includes the 90 days of follow-up excluded at the start.

Notes on coding: The analyses are done using a loop to analyze all outcomes for all contrasts in one go. The "cohort names" are:
- data1: novel GLP1 (sema + tirze) vs SGLT2
- data2: novel GLP1 vs DPP4
- data3: nvel GLP1 vs sulfo

**Instructions for sensitivity analyses:**
- Apply entropy balance weighting by changing iptw -> ebal (lines 234 and 365)
- Apply IPCW weights to per-protocol analysis by changing iptw to swfinal (line 365)
- Require an additional 90 days of follow-up: filter to keep patients with at leat 90 days of follow-up in ITT and PP analyses (lines 169 and 302). In that analysis, patients who experience the outcome or are otherwise lost to follow-up within 6 months are excluded. 
- Include early events and patients with <90 days of follow-up: change follow-up to start on index date (line 146). That analysis still requires at least 1 day of follow-up.  

```{r}
#### LOAD DATA ####

project_root <- Sys.getenv("PROJECT_ROOT")
if (nchar(project_root) == 0) stop("PROJECT_ROOT is not set. Please run: source config.sh (bash) or set it in .Renviron (RStudio).")
output_dir  <- file.path(project_root, "output")
results_dir <- file.path(project_root, "novel_res")

library(dplyr)
library(magrittr)
library(table1)
library(lubridate)
library(survival)
library(survminer)
library(tidyr)
library(broom)

end_date <- ymd(20260101)

data <- readRDS(file.path(output_dir, "data_novel.rds"))
data1 <- readRDS(file.path(output_dir, "data1_novel.rds"))
data2 <- readRDS(file.path(output_dir, "data2_novel.rds"))
data3 <- readRDS(file.path(output_dir, "data3_novel.rds"))
cov_list <- readRDS(file.path(output_dir, "cov_list_novel.rds"))

max_fu <- 730 # truncate follow-up up to 2 years after entry

#### 1. ALL-CAUSE DEMENTIA & SUBTYPES ####

# here, patients will not be censored if they experience all-cause dementia
# since all-cause dementia or its subtypes are the events of interest

# we are also incorporating the negative control outcomes in this loop
# since similarly, we would not censor at all-cause dementia for those

# save results from all models
all_cox_results <- list()

# save follow-up times and event count for each analysis
followup_times <- data.frame(
  cohort = NA,
  approach = NA,
  outcome = NA,
  mean = NA,
  min = NA,
  max = NA,
  median = NA,
  q1 = NA,
  q3 = NA,
  n_trt = NA,
  n_comp = NA,
  n_events = NA,
  n_events_trt = NA,
  n_events_comp = NA
)

# save proportional hazards assumption test for each model
cox_ph <- data.frame(
  cohort = NA,
  approach = NA,
  outcome = NA,
  pvalue = NA
)

# define outcomes as named within dataset var names:
# ex: "dementia_allcause" with var names "outcome_dementia_allcause_date" and res will be saved as "dementia_allcause"

dementia_outcomes <- c(
  # dementia outcomes:
  "dementia_allcause",

  # negative control outcomes:
  "appendicitis",
  "tooth",
  "bell",
  "basal"
)

for (outcome in c(dementia_outcomes)) {
  
  for (cohort_name in c("data1", "data2", "data3")) {

    cohort <- get(cohort_name)
    
    date_name <- paste0("outcome_", outcome, "_date")
    
    
    if (cohort_name == "data1") {
      ref <- "SGLT2"
      index <- "Semaglutide/Tirzepatide"
    } else if (cohort_name == "data2") {
      ref <- "DPP4"
      index <- "Semaglutide/Tirzepatide"
    } else if (cohort_name == "data3") {
      ref <- "Sulfonylureas"
      index <- "Semaglutide/Tirzepatide"
    }

    
    ### ITT APPROACH ###
    
    # describe follow-up (ITT)
    cohort$study_end <- end_date
    
    cohort$itt_exit_date <- pmin(
      cohort$dod, # date of death
      cohort$cov_end, # end of coverage
      cohort$study_end, # end of study
      cohort$index_date + max_fu, 
      na.rm = TRUE
    )
    
    summary(cohort$itt_exit_date)

    # now we are adding 90 days to index date to start follow up at 90 days
    # and reduce the influence of early events (likely caused by confounding by indication/reverse causality)
    cohort$f_up_start <- cohort$index_date + 90
    
    # for sensitivity analysis where start follow up immediately: 
    # cohort$f_up_start <- cohort$index_date

    # create indicator for itt_event
    # no event if no date of event or if occurred after ITT exit
    # !!sym(date_name) takes the date for the event being analyzed in this loop

    cohort %<>%
      mutate (itt_event = if_else(!is.na(!!sym(date_name)) & !!sym(date_name) < itt_exit_date & !!sym(date_name) >= f_up_start, 1, 0),
              itt_event_date = if_else(itt_event == 1, !!sym(date_name), NA))
    
    table(cohort$itt_event)
    
    # update itt exit date if there was an event
    # so that follow-up additionally ends at event occurrence
    cohort %<>%
      mutate(itt_exit_date = if_else(itt_event == 1, itt_event_date, itt_exit_date))
    
    cohort$itt_follow_up <- time_length(interval(as.Date(cohort$f_up_start), as.Date(cohort$itt_exit_date)), 'days')
    summary(cohort$itt_follow_up)

    cohort %<>% filter(itt_follow_up>0) 
    
    # sensitivity analysis requiring an additional 90 days of follow-up: 
    # cohort %<>% filter(itt_follow_up>=90) 

    # fill in info on follow-up
    followup_times <- rbind(
      followup_times,
      data.frame(
        cohort = cohort_name,
        approach = "itt",
        outcome = outcome,
        mean = mean(cohort$itt_follow_up),
        min = min(cohort$itt_follow_up),
        max = max(cohort$itt_follow_up),
        median = median(cohort$itt_follow_up),
        q1 = quantile(cohort$itt_follow_up, 0.25),
        q3 = quantile(cohort$itt_follow_up, 0.75),
        n_trt = sum(cohort$trt==1),
        n_comp = sum(cohort$trt==0),
        n_events = sum(cohort$itt_event),
        n_events_trt = sum(cohort$itt_event[cohort$trt==1]),
        n_events_comp = sum(cohort$itt_event[cohort$trt==0])
      )
    )
    
    
    # fit survival models
    # crude ITT model: 
    cox_mod_itt_crude <- coxph(Surv(itt_follow_up, itt_event) ~ trt, 
                         data = cohort, 
                         cluster = PATIENT_ID, 
                         robust = TRUE)

    coxph_p <- cox.zph(cox_mod_itt_crude)$table["trt", "p"]
    
    cox_ph <- rbind(
      cox_ph,
      data.frame(
        cohort = cohort_name,
        approach = "itt crude",
        outcome = outcome,
        pvalue = coxph_p
      )
    )
    
    cox_mod_itt_res_crude <- tidy(cox_mod_itt_crude, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "itt_crude"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results[[paste(outcome, cohort_name, "itt_crude", sep = "_")]] <- cox_mod_itt_res_crude
    
    km_curve_itt_crude <- survfit(Surv(itt_follow_up, itt_event) ~ trt, 
                            data = cohort)
    
    
    
    
    # stabilized IPTW weighted model:
    cox_mod_itt <- coxph(Surv(itt_follow_up, itt_event) ~ trt, 
                         data = cohort, 
                         weights = siptw, # change to ebal for entropy balancing weights (sensitivity analysis)
                         cluster = PATIENT_ID, 
                         robust = TRUE)

    coxph_p <- cox.zph(cox_mod_itt)$table["trt", "p"]
    
    cox_ph <- rbind(
      cox_ph,
      data.frame(
        cohort = cohort_name,
        approach = "itt IPTW",
        outcome = outcome,
        pvalue = coxph_p
      )
    )

    cox_mod_itt_res <- tidy(cox_mod_itt, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "itt"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results[[paste(outcome, cohort_name, "itt", sep = "_")]] <- cox_mod_itt_res
    
    
    
    
    
    ### PER PROTOCOL APPROACH ###
    
    # describe follow-up (PP)
    # same as above, but additionally censoring at switch and disc
    cohort$pp_exit_date <- pmin(
      cohort$dod,
      cohort$cov_end,
      cohort$study_end,
      cohort$switch_date,
      cohort$disc_date,
      cohort$index_date + max_fu,
      na.rm = TRUE
    )
    
    summary(cohort$pp_exit_date)
    
    # create indicator for pp_event
    # no event if no date of event or if occurred after PP exit
    cohort %<>%
      mutate (pp_event = if_else(!is.na(!!sym(date_name)) & !!sym(date_name) < pp_exit_date & !!sym(date_name) >= f_up_start, 1, 0),
              pp_event_date = if_else(pp_event == 1, !!sym(date_name), NA))
    
    table(cohort$pp_event)
    
    # update pp exit date if there was an event
    # so that follow-up additionally ends at event occurrence
    cohort %<>%
      mutate(pp_exit_date = if_else(pp_event == 1, pp_event_date, pp_exit_date))
    
    
    cohort$pp_follow_up <- time_length(interval(as.Date(cohort$f_up_start), as.Date(cohort$pp_exit_date)), 'days')
    summary(cohort$pp_follow_up)

    cohort %<>% filter(pp_follow_up>0)

    # sensitivity analysis requiring an additional 90 days of follow-up: 
    # cohort %<>% filter(pp_follow_up>=90) 
  
    
    # fill in info on follow-up
    followup_times <- rbind(
      followup_times,
      data.frame(
        cohort = cohort_name,
        approach = "pp",
        outcome = outcome,
        mean = mean(cohort$pp_follow_up),
        min = min(cohort$pp_follow_up),
        max = max(cohort$pp_follow_up),
        median = median(cohort$pp_follow_up),
        q1 = quantile(cohort$pp_follow_up, 0.25),
        q3 = quantile(cohort$pp_follow_up, 0.75),
        n_trt = sum(cohort$trt==1),
        n_comp = sum(cohort$trt==0),
        n_events = sum(cohort$pp_event),
        n_events_trt = sum(cohort$pp_event[cohort$trt==1]),
        n_events_comp = sum(cohort$pp_event[cohort$trt==0])
      )
    )
    
    
    # fit survival models
    # crude model:
    cox_mod_pp_crude <- coxph(Surv(pp_follow_up, pp_event) ~ trt, 
                        data = cohort,
                        cluster = PATIENT_ID, 
                        robust = TRUE)

    coxph_p <- cox.zph(cox_mod_pp_crude)$table["trt", "p"]
    
    cox_ph <- rbind(
      cox_ph,
      data.frame(
        cohort = cohort_name,
        approach = "pp crude",
        outcome = outcome,
        pvalue = coxph_p
      )
    )
    
    cox_mod_pp_res_crude <- tidy(cox_mod_pp_crude, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "pp_crude"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results[[paste(outcome, cohort_name, "pp_crude", sep = "_")]] <- cox_mod_pp_res_crude
    

    
    # stabilized IPTW:
    cox_mod_pp <- coxph(Surv(pp_follow_up, pp_event) ~ trt, 
                            data = cohort, 
                            weights = siptw, # change to ebal (IPTW only) or ebalfinal (ebal + IPCW) for sensitivity analysis using entropy balancing weights
                            # and change to swfinal for IPTW + IPCW weighting
                            cluster = PATIENT_ID, 
                            robust = TRUE)

    coxph_p <- cox.zph(cox_mod_pp)$table["trt", "p"]
    
    cox_ph <- rbind(
      cox_ph,
      data.frame(
        cohort = cohort_name,
        approach = "pp IPTW+IPCW",
        outcome = outcome,
        pvalue = coxph_p
      )
    )
    
    cox_mod_pp_res <- tidy(cox_mod_pp, exponentiate = TRUE, conf.int = TRUE) %>%
      mutate(
        cohort = cohort_name,
        outcome = outcome,
        approach = "pp"
      ) %>%
      select(cohort, outcome, approach, term, 
             hr = estimate, ci_lower = conf.low, ci_upper = conf.high, 
             pvalue = p.value)
    
    
    all_cox_results[[paste(outcome, cohort_name, "pp", sep = "_")]] <- cox_mod_pp_res
    
    
  }
}


rm(cohort, cohort_name, outcome)


#### SAVE MODEL RESULTS ####

write.csv(
  followup_times,
  paste0(results_dir, "/followup_times_", Sys.Date(), ".csv"),
  row.names = FALSE
)

cox_ph %<>%
  mutate(
    flag = if_else(pvalue<0.05, 1, 0)
  )

write.csv(
  cox_ph,
  paste0(results_dir, "/cox_ph_", Sys.Date(), ".csv"),
  row.names = FALSE
)

summary_table <- bind_rows(all_cox_results)

summary_table <- summary_table %>%
  mutate(across(where(is.numeric), ~round(., 2)))


write.csv(
  summary_table,
  paste0(results_dir, "/all_cox_results_", Sys.Date(), ".csv"),
  row.names = FALSE
)

saveRDS(
  summary_table,
  paste0(results_dir, "/all_cox_results_novel_", Sys.Date(), ".rds")
)

```

## FINAL TABLE FOR MANUSCRIPT 

```{r}
#### SAVE FINAL TABLE 2 ####

final_tab2_data1 <- summary_table %>%
  filter(cohort == "data1") %>% # keep only results for GLP1 vs SGLT2
  filter(outcome == "dementia_allcause") %>%
  filter(approach == "itt" | approach == "pp") %>% # keep only adjusted estimates
  mutate(estimate = paste0(hr, " (", ci_lower, ", ", ci_upper, ")")) %>% # format estimate and CI into same cell
  select(approach, estimate)

final_tab2_data2 <- summary_table %>%
  filter(cohort == "data2") %>% # keep only results for GLP1 vs DPP4
  filter(outcome == "dementia_allcause") %>%
  filter(approach == "itt" | approach == "pp") %>% # keep only adjusted estimates
  mutate(estimate = paste0(hr, " (", ci_lower, ", ", ci_upper, ")")) %>% # format estimate and CI into same cell
  select(approach, estimate)

final_tab2_data3 <- summary_table %>%
  filter(cohort == "data3") %>% # keep only results for GLP1 vs sulfo
  filter(outcome == "dementia_allcause") %>%
  filter(approach == "itt" | approach == "pp") %>% # keep only adjusted estimates
  mutate(estimate = paste0(hr, " (", ci_lower, ", ", ci_upper, ")")) %>% # format estimate and CI into same cell
  select(approach, estimate)

final_tab2 <- data.frame(
  glp1_vs_sglt2 = c(final_tab2_data1$estimate[final_tab2_data1$approach=="itt"], final_tab2_data1$estimate[final_tab2_data1$approach=="pp"]),
  glp1_vs_dpp4 = c(final_tab2_data2$estimate[final_tab2_data2$approach=="itt"], final_tab2_data2$estimate[final_tab2_data2$approach=="pp"]),
  glp1_vs_sulfo = c(final_tab2_data3$estimate[final_tab2_data3$approach=="itt"], final_tab2_data3$estimate[final_tab2_data3$approach=="pp"])
)

row.names(final_tab2) <- c("ITT", "PP")

write.csv(
  final_tab2,
  paste0(results_dir, "/final_tab2_", Sys.Date(), ".csv"),
  row.names = FALSE
)
```
